---
output:
  pdf_document: default
  html_notebook: default
---

# Duncan McKinnon
# West
# W 1

## QA

Load Packages and Data

```{r message=FALSE, warning=FALSE}

suppressPackageStartupMessages({
  library(tidyverse)
  library(okcupiddata)
})

```

### 1).

```{r}
# get top ten cities by size

locations <- profiles %>% 
  group_by(location) %>% 
  summarize(num = n()) %>% 
  separate(location, into = c('city', 'state'), sep = ', ', remove = TRUE) %>%
  arrange(desc(num))

top_ten <- locations$city[1:10]
top_ten
```

### 2).

```{r}
ggplot(locations %>% filter(num > 2)) + 
  geom_col(aes(x = city, y = num)) +
  scale_x_discrete(breaks = top_ten, labels = top_ten) +
  labs(title = 'distribution of okcupid data users', 
       x = 'city (top ten labeled)', 
       y = 'count') +
  theme(axis.text.x = element_text(angle = 67.5, hjust = 1)) 
```

### 3).

```{r}

ggplot(profiles %>% filter(!is.na(age))) +
  geom_histogram(aes(x = age), bins = 25) + 
  labs(title = 'ages of okcupid data users', 
       x = 'age', 
       y = 'count') +
  theme_bw()
```

### 4).

```{r}

ggplot(profiles %>% filter(!is.na(height))) +
  geom_histogram(aes(x = height), bins = 25) + 
  labs(title = 'height of okcupid data users', 
       x = 'height', 
       y = 'count') +
  theme_bw()
```

### 5).

Height in this dataset is very normally distributed with a mean around 72 inches (6 ft).  While this mean is above what would be expected for a mixed population of men and women we could probably expect people to give optimistic heights up to at least 6ft.

```{r}
ggplot(profiles %>% filter(!is.na(income))) +
  geom_histogram(aes(x = income), binwidth = 10000) + 
  labs(title = 'income of okcupid data users', 
       x = 'income', 
       y = 'count') +
  theme_bw()
```

### 6).

Income was pretty sparse in this dataset, with the majority of participants likely opting out.  The distribution of incomes is very right skewed, with a high concentration between 20k and 100k and then smaller peaks around 250k, 500k.  The right skew is again effected by the hard lower limit of $0 income.  The maximum entry of 1m is a lot more frequent than 250k or 500k, so it seems probable that there was also a hard upper limit in the question, or even a set of categories to choose from.
